---
title: "RestHabitsMD"
author: "Stella & Lauren"
date: "11/16/2021"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### BEGIN PROJECT

**load packages**

```{r load packages}
library(tidyverse)
library(knitr)
library(ggmosaic)
library(stringr)
library(wordcloud2)
```

**reading in data**

```{r read data}
rest_habits_raw <- readr::read_csv(here::here("data","418FinalProjData.csv"))
```

### CLEAN DATA

**Goals:**

- delete the extra columns
- do we need to rename?
- check for missing/incorrect values (what do we do?)

Rundown of variables:

| Name (Rename)                  |  Description
|--------------------------------|--------------------------------
|`age` [convert to num]          | ages 18+
|`sex`                           | Male/Female/Other
|`race``                         | Race/ethnicity
|`year`                          | school year (1-4, 5+)
|`gpa_1` (gpa[convert to num])   | (should be in 0.00 format) (check?)

|`degree`                        | undergraduate/graduate/combined degree
|`undergrad`                     | 
|`grad`                          | 
|`combined`                      | 

|`sleepDuration`                 | duration slept each night
|`sleepEnough`                   | is sleepDuration enough? (Y/N)

|`regSleep`                      | do they have a regular sleep schedule? (Y/N)
|`sleepCatchUp`                  | "catch-up" on sleep? (Y/N)
|`allNighter`                    | pull all nighers?
|`dayTired`                      | feel tired during the day?

|`fallStayAsleep`                | have issues falling/staying asleep?
|`sleepDisorder`                 | previously diagnosed with a sleep disorder?

|`napFreq`                       | how frequently do they nap?
|`napDuration`                   | how long are the naps?

|`leisureTime`                   | have leisure time?
|`leisureFreq`                   | how much leisure time?

|`socialMedia`                   | do they use social media?
|`socMedDuration`                | how much time do they spend on social media?

|`academic_1` (academics)        | how do they feel they are doing academically?
|`mentalHealth_1` (energy)       | energy levels
|`mentalHealth_2` (stress)       | stress levels
|`mentalHealth_3` (concentration)| ability to concentrate
|`mentalHealth_4` (mood)         | overall mood

- delete unwanted cols/rows
  
```{r remove cols}
rest_habits <- rest_habits_raw %>% 
  select(age:`mentalHealth _4`)
```

- removed first 2 rows & blank observations

```{r remove first row}
rest_habits <- rest_habits [-c(1,2),]

rest_habits <- rest_habits %>% 
  filter(!is.na(sex))
```

- renaming

```{r rename}
rest_habits <- rest_habits %>% 
  rename(gpa=`gpa _1`,
         academics=`academic _1`,
         energy=`mentalHealth _1`,
         stress=`mentalHealth _2`,
         concentration=`mentalHealth _3`,
         mood=`mentalHealth _4`)
```

- convert age & gpa to numeric

```{r num gpa/age}
rest_habits$gpa <- as.numeric(rest_habits$gpa)
rest_habits$age <- as.numeric(rest_habits$age)
```


```{r making category column}
rest_habits$category <- rest_habits %>%
    add_column(category = NA)
```


```{r group majors into category}
rest_habits <- rest_habits %>%
   mutate(category = case_when(
   #Business & Legal
   grad %in% c("Communications", "Business Administration", "Criminal Justice") ~ "businessLegal",
   #Cultures and Humanities
   grad %in% c("") ~ "humanities",
   #Biological, Environmental, and Physical Sciences
   grad %in% c("Physics") ~ "science",
   #Education
   grad %in% c("Education", "Higher Education", "School Psychology") ~ "education",
   #Fine Arts and Design
   grad %in% ("") ~ "art",
   #Health and Medicine
   grad %in% c("Physician Assistant Studies", "Medical Dosimetry", "Public Health", "Physical Therapy", "Nursing") ~ "health",
   #Math and Engineering
   grad %in% c("Biostatistics", "Data Science and Analytics") ~ "mathEngineering",
   #Social and Behavioral Sciences
   grad %in% c("Speech-Language Pathology", "Social Work") ~ "social",

    undergrad %in% c("Multimedia Journalism", "Marketing", "Accounting", "Finance", "International Business", "Advertising and Public Relations", "Business, General", "Management", "Hospitality and Tourism Management", "Human Resource Management", "Communication Studies", "Criminal Justice", "Legal Studies", "Supply Chain Management") ~ "businessLegal",
    undergrad %in% c("French", "Spanish", "History", "English", "Writing", "Philosophy") ~ "humanities",
    undergrad %in% c("Biology", "Biochemistry", "Cell and Molecular Biology", "Wildlife Biology", "Natural Resources Management", "Environmental and Sustainability Studies", "Applied Food and Nutrition")  ~ "science",
    undergrad %in% c("Art Education") ~ "education",
    undergrad %in% c("Film and Video") ~ "art",
    undergrad %in% c("Physician Assistant Studies", "Medical Dosimetry", "Public Health", "Physical Therapy", "Nursing", "Preprofessional Studies", "Exercise Science", "Allied Health Sciences", "Biomedical Sciences", "Radiation Therapy", "Recreational Therapy", "Health Information Management") ~ "health",
    undergrad %in% c("Electrical Engineering", "Biostatistics", "Statistics", "Mathematics", "Biomedical Engineering", "Computer Science", "Cybersecurity") ~ "mathEngineering",
    undergrad %in% c("Speech-Language Pathology", "Social Work", "Psychology") ~ "social",

     combined %in% c("Accounting") ~ "businessLegal",
     combined %in% c("Cell and Molecular Biology") ~ "science",
   combined %in% c("Exercise Science and Athletic Training") ~ "health",
     combined %in% c("Statistics", "Biostatistics", "Cybersecurity", "Statistics and Biostatistics") ~ "mathEngineering"
     )
   )
```


```{r define levels}
levels_sleepDur = c('Less than 5 hours.', '5-6 hours.', '6-7 hours.', '7-8 hours.', '9-10 hours.', 'More than 10 hours.')

levels_acad = c('Bad', 'Not Good', 'Ok', 'Good', 'Great') # for acad & mental health

levels_socMedDur = c('Less than 1 hour/day.', '1-3 hours/day.', '3-6 hours/day.', '6-9 hours/day.') # for duration of Social Media

levels_allNighter = c('No, never.', 'Yes, rarely.', 'Yes, I do it often.') # for all nighter

levels_leisureFreq = c('I am too busy with school/work to have regular leisure time in my schedule.', 'I spend some time on personal activities a few times a week.', '1-3 hours/day.', '3-6 hours/day.')
```


```{r factor levels}
rest_habits <- rest_habits %>% 
  mutate(sleepDuration = factor(sleepDuration, levels = levels_sleepDur)) %>% 
  mutate(across(c(academics, concentration, mood, stress, energy), ~factor(., levels = levels_acad))) %>% 
  mutate(socMedDuration = factor(socMedDuration, levels = levels_socMedDur)) %>% 
  mutate(allNighter = factor(allNighter, levels = levels_allNighter)) %>% 
  mutate(leisureFreq = factor(leisureFreq, levels = levels_leisureFreq))
```

*writing out cleaned dataset for shiny app*

```{r output for shiny}
readr::write_rds(rest_habits, here::here("RestHabits/data", "shinydata.rds"))
```


### MESSING WITH VISUALIZATIONS

```{r gpa histogram}
rest_habits %>% 
  filter(!is.na(gpa)) %>% 
  ggplot(aes(x=gpa)) + 
  geom_histogram(fill="lightcoral") +
  geom_vline(aes(xintercept = mean(gpa)), linetype="dashed")
```

**TO DO** what vars do we want 4 realsies?

```{r contingency table}
c_table <-
  rest_habits %>% 
  group_by(sex, sleepDisorder) %>% 
  tally() %>% 
  spread(key = sleepDisorder, value = n)

print.data.frame(c_table)
```

(trying contingency table but *kable*)

```{r contingency kable}
rest_habits %>%
  group_by(sex, sleepDisorder) %>%
  summarize(n=n()) %>%
  spread(sleepDisorder, n)%>%
  kable()
```


```{r mosaic}
rest_habits %>% 
  filter(!is.na(sleepDuration)) %>% 
  ggplot() +
  geom_mosaic(aes(x=product(sleepDuration, degree), fill=sleepDuration)) +
  labs(y="Duration of sleep each night", x="Type of Degree", title="Mosaic Plot")
```

 - basic boxplot? even need or just stick to hist?

```{r boxplot gpa}
rest_habits %>% 
  filter(!is.na(gpa)) %>% 
  filter(!is.na(leisureTime)) %>% 
  ggplot(aes(x=leisureTime, y=gpa, fill=leisureTime)) + 
  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=20, size=8, color="grey", fill="grey")
```


```{r bar chart degree}
ggplot(rest_habits, aes(x=degree)) + 
  geom_bar()
```


```{r bar chart concentration}
rest_habits %>% 
  ggplot(aes(x=concentration)) + 
  geom_bar() +
  coord_flip()
```


```{r bar chart energy vs sleep dur.}
 rest_habits %>% 
  filter(!is.na(sleepDuration)) %>%
  filter(!is.na(energy)) %>%
            ggplot(aes(x = energy, fill = sleepDuration)) + 
            geom_bar(position = "dodge")
```
```{r lt vs soc med}
rest_habits %>% 
  filter(!is.na(socMedDuration)) %>% 
  ggplot() +
  geom_mosaic(aes(x=product(leisureTime, socMedDuration), fill=leisureTime)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
rest_habits %>% 
  filter(!is.na(socMedDuration)) %>% 
  filter(!is.na(leisureFreq)) %>% 
  mutate(leisureFreq = str_wrap(leisureFreq, width = 20)) %>% 
  ggplot() +
  geom_mosaic(aes(x=product(leisureFreq, socMedDuration), fill=leisureFreq)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
rest_habits %>% 
  filter(!is.na(category)) %>% 
  group_by(category) %>% 
  summarize(count = n(),
            wordFreq = count/129) %>% 
  wordcloud2(size=.5)
```


```{r sleep vs major}
rest_habits %>% 
  filter(!is.na(category)) %>% 
  filter(!is.na(sleepDuration)) %>% 
  ggplot() +
  geom_mosaic(aes(x=product(sleepDuration, category), fill=sleepDuration)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r gpa vs major}
rest_habits %>% 
  filter(!is.na(gpa)) %>% 
  filter(!is.na(category)) %>% 
  ggplot(aes(x=category, y=gpa, fill=category)) + 
  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=20, size=4, color="grey", fill="grey") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


### QUESTIONS

- do particular degrees/majors sleep less than others?

- do those who sleep less have worse mental health/academics?

- do those who nap tend to have higher gpa/academic rating?

- do students think that the amount they sleep each night is enough?

- do students who pull all-nighters have a lower gpa/academics?

- which majors have less leisure time?

- leisure time vs. gpa?

- gpa vs. academics?

- no leisure time vs. time on social media?

- nap freq/duration vs. energy?


### SOURCES

contingency tables: https://community.rstudio.com/t/how-do-you-convert-a-tidy-data-frame-to-a-contingency-table/99562/2

http://analyticswithr.com/contingencytables.html

box plots: http://www.sthda.com/english/wiki/ggplot2-box-plot-quick-start-guide-r-software-and-data-visualization

bar charts: http://www.sthda.com/english/wiki/ggplot2-barplots-quick-start-guide-r-software-and-data-visualization

mosaic plots: https://cran.r-project.org/web/packages/ggmosaic/vignettes/ggmosaic.html

circular bar plot: https://www.r-graph-gallery.com/web-circular-barplot-with-R-and-ggplot2.html looks cool af

analysis? https://scc.ms.unimelb.edu.au/resources-list/simple-r-scripts-for-analysis/r-scripts

themes https://ggplot2.tidyverse.org/reference/ggtheme.html



data <- tibble (
  q1 = sample(n, c(...)) run simulations to make fake 
)
